---
title: "Gov 50 Final Project"
author: "Olivia Data"
description: "My Final Project"
output:
  distill::distill_article:
    self_contained: false
---

What effects, if any, does party alignment have on American's support of voting rights? My hypothesis is that Democrats will be more likely than Republicans or Independents to support expanding voting rights and making it easier for more people to vote. However, I am not sure if a correlation between supporting voting rights and being a Democrat would be causal, as factors such as race, income, and education would also strongly affect a person's support of expanding voting rights. I believe race (because voting restrictions often disproportionaly impact people of color) may be a greater factor than party.

My explanatory variables in this question are party alignment as well as several demographic factors. The data I've found from the Pew Research center (https://www.pewresearch.org/politics/2021/04/22/republicans-and-democrats-move-further-apart-in-views-of-voting-access/) is from a survey, so these variables were measured by how respondents reported their identities and opinions.

The outcome variable is whether or not a person supports expanding voting rights. This was also measured through a survey in which the options were strongly support, somewhat support, somewhat disapprove, or strongly disapprove. There were a number of questions regarding different types of voting restrictions and expansions.

If my hypothesis is correct, then a higher percentage of Democrats would say they approve of expanding voting rights than Republicans or Independents. If my hypothesis is incorrect, then these percentages will either be roughly equivalent, or a lower percentage of Democrats would support the expansion of various voting rights.



```{r}
library(tidyverse)
library(readr)
library(haven)
data <- read_sav("W87_Apr21/ATP W87.sav")

data
```
```{r}
library(ggplot2)
data |>
  mutate(party = case_when(F_PARTY_FINAL == 1 ~ "Republican",
                          F_PARTY_FINAL == 2 ~ "Democrat",
                          F_PARTY_FINAL == 3 ~ "Independent",
                          F_PARTY_FINAL == 4 ~ "Something Else"),
         voting = case_when(VTPRIORITY_e_W87 == 1 ~ "Strongly Favor",
                          VTPRIORITY_e_W87 == 2 ~ "Somewhat Favor",
                          VTPRIORITY_e_W87 == 3 ~ "Somewhat Oppose",
                          VTPRIORITY_e_W87 == 4 ~ "Strongly Oppose",
                          VTPRIORITY_e_W87 == 5 ~ "No Answer")) |>
   ggplot(mapping = aes(x = party,
                       y = voting)) +
  geom_count() +
  labs(x = "Party",
       y = "Approval of Allowing People Convicted of Felonies to Vote After Serving Their Sentences",
       title = "Voting Restriction Opinions by Party")
```
In this visualization, it is clear that Democrats seem to be the most supportive of allowing people convicted of a felony to vote after serving their time, and Republicans seem more likely to oppose allowing them to vote.

```{r}
library(ggplot2)
data |>
  mutate(race = case_when(F_RACECMB == 1 ~ "White",
                          F_RACECMB == 2 ~ "Black",
                          F_RACECMB == 3 ~ "Asian/Asian-American",
                          F_RACECMB == 4 ~ "Mixed Race",
                          F_RACECMB == 5 ~ "Other",
                          F_RACECMB == 6 ~ "Don’t know/Refused",),
         voting = case_when(VTPRIORITY_e_W87 == 1 ~ "Strongly Favor",
                          VTPRIORITY_e_W87 == 2 ~ "Somewhat Favor",
                          VTPRIORITY_e_W87 == 3 ~ "Somewhat Oppose",
                          VTPRIORITY_e_W87 == 4 ~ "Strongly Oppose",
                          VTPRIORITY_e_W87 == 5 ~ "No Answer")) |>
   ggplot(mapping = aes(x = race,
                       y = voting)) +
  geom_count() +
  labs(x = "Race",
       y = "Approval of Allowing People Convicted of Felonies to Vote After Serving Their Sentences",
       title = "Voting Restriction Opinions by Race")
```

This visualization shows the breakdown of opinions on allowing people convicted of felonies to vote after serving their sentence by race. Race could be a confounding variable in whether or not party alignment affects people's opinions of voting rights, Black people often lean Democrat and often support greater expansions in voting rights. However, it seems in this chart that White people, Asian and Asian-American people, and other races are also more often supportive of expanding voting rights.

