---
title: "Gov 50 Final Project"
author: "Olivia Data"
description: "Political Parties and Voting Rights"
output:
  distill::distill_article:
    self_contained: false
---
Introduction:

What effects, if any, does party alignment have on whether or not Americans support voting rights for people convicted of felonies? My hypothesis is that Democrats will be more likely than Republicans or Independents to support allowing people convicted of felonies to vote after serving their sentence. However, I am not sure if a correlation between supporting voting rights and being a Democrat would be causal, as factors such as race, income, and education would also most likely have a strong effect on a person's support of expanding such voting rights. After all, voting restrictions often disproportionately impact people of color. Also, people from lower income areas as well as people of color are more likely to be targeted by the police and more likely not to have access to resources such as housing or therapy that might reduce a person's likelihood of being involved in crime. I believe that party alignment will be a greater factor in whether or not a person supports allowing people convicted of felonies to vote after serving their sentence because there are bound to be differences of opinions in any given demographic, but there are thus several confounding factors to consider.

My Data:

The data I've found from the Pew Research center (https://www.pewresearch.org/politics/2021/04/22/republicans-and-democrats-move-further-apart-in-views-of-voting-access/) is from a survey, so these variables were measured by how respondents reported their identities and opinions. The surveys were sent to a random sample of households based on addresses randomly selected from the U.S Postal Service's Delivery Sequence File. This research design is cross-sectional.

My explanatory variable in this question is party alignment. I also compare party alignment to demographic factors such as race, income, or education level, with a particular focus on race. 

The outcome variable is whether or not a respondent supports allowing people convicted of felonies to vote after serving their sentence. This was also measured through a survey in which the options were strongly support, somewhat support, somewhat disapprove, or strongly disapprove.

If my hypothesis is correct, then a higher percentage of Democrats would say they approve of felons' voting rights than Republicans or Independents. If my hypothesis is incorrect, then these percentages will either be roughly equivalent, or a lower percentage of Democrats would support the expansion of various voting rights.


```{r}
library(tidyverse)
library(readr)
library(haven)
data <- read_sav("W87_Apr21/ATP W87.sav")

data
```
```{r}
library(ggplot2)
data |>
  mutate(voting = case_when(VTPRIORITY_e_W87 == 1 ~ "Strongly Favor",
                          VTPRIORITY_e_W87 == 2 ~ "Somewhat Favor",
                          VTPRIORITY_e_W87 == 3 ~ "Somewhat Oppose",
                          VTPRIORITY_e_W87 == 4 ~ "Strongly Oppose",
                          VTPRIORITY_e_W87 == 99 ~ "No Answer")) |>
   ggplot(mapping = aes(x = voting)) +
  geom_bar() +
  labs(x = "Opinion on Allowing People Convicted of a Felony to Vote After Serving Their Sentence",
       y = "Count",
       title = "Distribution of Dependent Variable")
```
The above graph shows the distribution of the dependent variable, which is people's opinion on whether or not people convicted of felonies should be allowed to vote after serving their sentence. It seems that the most popular opinion is to strongly favor allowing them to vote, followed by somewhat favoring allowing them to vote.

Results:

```{r}
library(ggplot2)
data |>
  mutate(party = case_when(F_PARTY_FINAL == 1 ~ "Republican",
                          F_PARTY_FINAL == 2 ~ "Democrat",
                          F_PARTY_FINAL == 3 ~ "Independent",
                          F_PARTY_FINAL == 4 ~ "Something Else",
                          F_PARTY_FINAL == 99 ~ "Refused to Answer"),
         voting = case_when(VTPRIORITY_e_W87 == 1 ~ "Strongly Favor",
                          VTPRIORITY_e_W87 == 2 ~ "Somewhat Favor",
                          VTPRIORITY_e_W87 == 3 ~ "Somewhat Oppose",
                          VTPRIORITY_e_W87 == 4 ~ "Strongly Oppose",
                          VTPRIORITY_e_W87 == 99 ~ "No Answer")) |>
   ggplot(mapping = aes(x = party,
                       y = voting)) +
  geom_count() +
  labs(x = "Party",
       y = "Approval of Allowing People Convicted of Felonies to Vote After Serving Their Sentences",
       title = "Voting Restriction Opinions by Party")
```
In this visualization, it is clear that Democrats seem to be the most supportive political party of allowing people convicted of a felony to vote after serving their time, and Republicans and Independents seem slightly more likely than Democrats to oppose allowing them to vote.

```{r}
library(ggplot2)
data |>
 mutate(race = case_when(F_RACECMB == 1 ~ "White",
                          F_RACECMB == 2 ~ "Black",
                          F_RACECMB == 3 ~ "Asian/Asian-American",
                          F_RACECMB == 4 ~ "Mixed Race",
                          F_RACECMB == 5 ~ "Other",
                          F_RACECMB == 99 ~ NA_character_),
         voting = case_when(VTPRIORITY_e_W87 == 1 ~ "Strongly Favor",
                          VTPRIORITY_e_W87 == 2 ~ "Somewhat Favor",
                          VTPRIORITY_e_W87 == 3 ~ "Somewhat Oppose",
                          VTPRIORITY_e_W87 == 4 ~ "Strongly Oppose",
                          VTPRIORITY_e_W87 == 99 ~ "No Answer")) |>
   ggplot(mapping = aes(x = race,
                       y = voting)) +
  geom_count() +
  labs(x = "Race",
       y = "Approval of Allowing People Convicted of Felonies to Vote After Serving Their Sentences",
       title = "Voting Restriction Opinions by Race")
```
This visualization shows the breakdown of opinions on allowing people convicted of felonies to vote after serving their sentence by race. Race could be a confounding variable in whether or not party alignment affects people's opinions of voting rights; Black people often lean Democrat and often support greater expansions in voting rights. However, it seems in this chart that White people, Asian and Asian-American people, and other races are also often supportive of felons' voting rights.


```{r}
race_data <- data |>
  mutate(race = case_when(F_RACECMB == 1 ~ "White",
                          F_RACECMB == 2 ~ "Black",
                          F_RACECMB == 3 ~ "Asian/Asian-American",
                          F_RACECMB == 4 ~ "Mixed Race",
                          F_RACECMB == 5 ~ "Other",
                          F_RACECMB == 99 ~ NA_character_),
         voting = case_when(as.numeric(VTPRIORITY_e_W87) == 99 ~ NA_real_,
                            TRUE ~ as.numeric(VTPRIORITY_e_W87)))

data_table <- race_data |>
  group_by(race) |>
  summarize(voting_mean = mean(voting, na.rm = TRUE)) |>
  pivot_wider(
    names_from = race, 
    values_from = voting_mean)

knitr::kable(data_table)
##x2 = recode(x, ‘1=4; 2=3; 3=2; 4=1’)
```
The above table shows the distribution of mean opposition of convicted felons voting after serving their sentences based on the race of the respondent. White people have the highest mean at about 2.07, while Black people have the lowest mean at about 1.483. This means that Black people, on average, show the highest level of support for allowing people convicted of felonies to vote after serving their sentence while White people, on average, show a lower level of support. 

```{r}
party_data <- data |>
  mutate(party = case_when(F_PARTY_FINAL == 1 ~ "Republican",
                          F_PARTY_FINAL == 2 ~ "Democrat",
                          F_PARTY_FINAL == 3 ~ NA_character_,
                          F_PARTY_FINAL == 4 ~ NA_character_,
                          F_PARTY_FINAL == 99 ~ NA_character_),
         voting = case_when(as.numeric(VTPRIORITY_e_W87) == 99 ~ NA_real_,
                            TRUE ~ as.numeric(VTPRIORITY_e_W87)))

party_table <- party_data |>
  group_by(party) |>
  summarize(voting_mean = mean(voting, na.rm = TRUE)) |>
  pivot_wider(
    names_from = party, 
    values_from = voting_mean) |>
  mutate(ATE = Democrat - Republican)
party_table
```
The above table shows the average treatment effect of being a Democrat on whether or not someone supports allowing people convicted of a felony to vote after serving their sentence. The average treatment effect is about -0.973, meaning that if someone is a Democrat, they are about 1 point more supportive on average of allowing people convicted of felonies to vote than if they are a Republican.

```{r}
library(infer)
data_filter <- data |>
  mutate(party = case_when(F_PARTY_FINAL == 1 ~ "Republican",
                          F_PARTY_FINAL == 2 ~ "Democrat",
                          F_PARTY_FINAL == 3 ~ "Independent",
                          F_PARTY_FINAL == 4 ~ "Something Else",
                          F_PARTY_FINAL == 6 ~ "Refused to Answer"),
         voting = case_when(as.numeric(VTPRIORITY_e_W87) <= 2 ~ "Favor",
                          as.numeric(VTPRIORITY_e_W87) >= 3 ~ "Oppose or No Answer")) |>
  filter(party%in% c("Republican", "Democrat"))

base_diff_attend <- data_filter |>
  infer::specify(voting ~ party, success = "Favor") |>
  infer::calculate(stat = "diff in props", order = c("Republican", "Democrat"))
base_diff_attend

base_p_attend <- data_filter |>
  specify(voting ~ party, success = "Favor") |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in props", order = c("Republican", "Democrat")) |>
  get_p_value(obs_stat = base_diff_attend, direction = "both")
base_p_attend
```
The p-value for the test of the null hypothesis that there is no difference in support of allowing people convicted of felonies to vote after serving their sentence between different political parties is about 0, meaning that there is a very small probability of getting an estimated difference as large as the one estimated by chance. At a level of alpha = 0.05, we can thus reject the null hypothesis and say that the effect of party on support of felons' voting is statistically significant.


```{r}
library(labelled)

fit <- lm(VTPRIORITY_e_W87 ~ unlabelled(F_PARTY_FINAL) + unlabelled(F_RACECMB) + unlabelled(F_EDUCCAT) + unlabelled(F_INC_TIER2), data = data)

prop_coef <- coef(fit)["prop"]

summary(fit)

```
The estimated coefficient on the prop for Democrats is -1.138 The p-value on this estimate is 0.00237, which is less than an alpha value of 0.05. Thus, at this level, we can reject the null hypothesis that party has no effect on opinion of felons voting. Even when accounting for other coefficients, like race, income, or education, this test is statistically significant.

Because of this multiple regression and the p-value calculated earlier, it appears that the relationship between party and opinion of felons' voting is likely a causal relationship.


Conclusion:

Overall, these results support my initial hypothesis that whether or not a respondent was a person of color would likely seem to have an effect on their support of felons' voting rights, but their political party would still likely have a causal relationship with support of felon's voting rights even when controlling for other factors. Indeed, the tables and calculated p-values show that Democrats are the most likely to believe that people convicted of felonies should be able to vote after serving their sentence. Race also is a factor in support, with Black people being the most likely to support these voting rights, but the multiple regression leads to the conclusion that party is a significant factor in support even when controlling for race.

One limitation with this experiment is its nature as a survey. People who self-report their identities and opinions may report incorrect information depending on what they think they should say rather than what they actually believe. While it might be difficult to fix such an error, future studies on this could always widen the sample size in order to account for this error and to help represent the population better.
